---
title: "MALTEM and schools"
subtitle: "A brief overview"
author: "Laia Cirera, Joe Brew, Judit Valls, Elisa Sicuri"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
fig_height: 5
fig_width: 7
output: ioslides_presentation

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r, results = 'hide'}
# Packages
library(ggplot2)
library(cism)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(leaflet)
library(readr)
library(ggthemes)
```


# Study area

```{r}
cols <- c('darkblue', 'darkgreen')
moz2_fortified <- moz2_fortified
map <- moz2_fortified %>%
  filter(id %in% c('Manhiça',
                   'Magude'))
ggplot() +
  geom_polygon(data = map,
               aes(x = long,
                   y = lat,
                   group = id,
                   fill = id),
               color = 'white',
               # color = 'grey',
               alpha = 0.8) +
  theme_map() +
  scale_fill_manual(name = '',
                    values = cols) +
  # geom_point(data = geo,
  #            aes(x = lng,
  #                y = lat),
  #            color = 'black',
  #            alpha = 0.7,
  #            size = 3) +
  # theme_black(base_size = 0) +
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank(),
        legend.position = 'none') +
  theme_cism_map()
```


```{r}
source('../prepare_student_data.R', chdir = TRUE)
ab <- ab %>% filter(year >= 2015)
performance <- performance %>% filter(year >= 2015)
ab <- ab %>% filter(!grepl('NA', year_term))
performance <- performance %>% filter(!grepl('NA', trimester))

library(RColorBrewer)
# Define some colors
cols <- colorRampPalette(brewer.pal(n = 9,
                                    name = 'BrBG'))(2)
tablify <- function(x){
  DT::datatable(x,
                selection = 'none',
                escape = FALSE,
                options = list(#sDom  = '<"top">lrt<"bottom">ip',
                               pageLength = 5, 
                               dom = 'tip'))
}
```

# Performance

## Number of unique students, per year

```{r}
x <- performance %>%
  group_by(year, name) %>%
  tally %>%
  ungroup %>%
  dplyr::select(-n) %>%
  filter(year >= 2015) %>%
  group_by(year) %>%
  tally 
names(x) <- c('Year', 'Students')
tablify(x)
```


## Number of unique students, per year, per district

```{r}
x <- performance %>%
  group_by(year, district, name) %>%
  tally %>%
  ungroup %>%
  dplyr::select(-n) %>%
  filter(year >= 2015) %>%
  group_by(year, district) %>%
  tally 
names(x) <- c('Year', 'District', 'Students')
tablify(x)
```


## Number of students followed through both years

```{r}
x <- performance %>%
  group_by(name, year) %>%
  tally %>%
  ungroup 
starters <- length(unique(x$name[x$year == 2015]))
finishers <- x %>%
  group_by(name) %>%
  filter(length(year) > 1) %>%
  dplyr::select(-n) %>%
  tally
finishers <- nrow(finishers)
```

`r finishers` of `r starters` (`r round(finishers / starters * 100, digits = 2)`%)

## Average number of observations, per student, per year

(One observation = 1 trimester-class)

```{r}
x <- 
  performance %>%
  group_by(name, year) %>%
  tally %>%
  ungroup %>%
  filter(year >= 2015) %>%
  group_by(year) %>%
  summarise(`Average Observations` = round(mean(n, digits = 2)))
tablify(x)
```

## Total number of observations, per year	

(One observation = 1 trimester-class)

```{r}
x <- 
  performance %>%
  group_by(name, year) %>%
  tally %>%
  ungroup %>%
  filter(year >= 2015) %>%
  group_by(year) %>%
  summarise(`Total Observations` = sum(n))
tablify(x)
```

## Average yearly grade, per school and year

```{r}
x <- 
  performance %>%
  filter(year >= 2015) %>%
  group_by(school, year) %>%
  summarise(`Average grade` = round(mean(value), digits = 2)) %>%
  ungroup %>%
  spread(key = year,
         value = `Average grade`)
tablify(x)
```
	
## Maps on performance	

Average grade by year

```{r}
cols <- c('darkblue', 'darkgreen')
moz2_fortified <- moz2_fortified
map <- moz2_fortified %>%
  filter(id %in% c('Manhiça',
                   'Magude'))

x <- 
  performance %>%
  filter(year >= 2015) %>%
  group_by(school, year) %>%
  summarise(`Average grade` = round(mean(value), digits = 2)) %>%
  ungroup %>%
  left_join(geo)

ggplot() +
  geom_polygon(data = map,
               aes(x = long,
                   y = lat,
                   group = id),
               color = 'white',
               alpha = 0.8) +
  geom_point(data = x,
             aes(x = lng,
                 y = lat,
                 color = `Average grade`),
             alpha = 0.7,
             size = 5) +
  facet_wrap(~year) +
  theme_cism_map() +
  scale_color_gradient(name = '',
                       low = 'red',
                       high = 'green')
```

# Absenteeism


## Number of unique students, per year

```{r}
x <- ab %>%
  group_by(year, name) %>%
  tally %>%
  ungroup %>%
  dplyr::select(-n) %>%
  filter(year >= 2015) %>%
  group_by(year) %>%
  tally 
names(x) <- c('Year', 'Students')
tablify(x)
```


## Number of unique students, per year, per district

```{r}
x <- ab %>%
  group_by(year, district, name) %>%
  tally %>%
  ungroup %>%
  dplyr::select(-n) %>%
  filter(year >= 2015) %>%
  group_by(year, district) %>%
  tally 
names(x) <- c('Year', 'District', 'Students')
tablify(x)
```


## Number of students followed through both years

```{r}
x <- ab %>%
  group_by(name, year) %>%
  tally %>%
  ungroup 
starters <- length(unique(x$name[x$year == 2015]))
finishers <- x %>%
  group_by(name) %>%
  filter(length(year) > 1) %>%
  dplyr::select(-n) %>%
  tally
finishers <- nrow(finishers)
```

`r finishers` of `r starters` (`r round(finishers / starters * 100, digits = 2)`%)

## Average number of observations, per student, per year

(One observation = 1 student-day)

```{r}
x <- 
  ab %>%
  group_by(name, year) %>%
  tally %>%
  ungroup %>%
  filter(year >= 2015) %>%
  group_by(year) %>%
  summarise(`Average Observations` = round(mean(n, digits = 2)))
tablify(x)
```

## Total number of observations, per year	

(One observation = 1 student-day)

```{r}
x <- 
  ab %>%
  group_by(name, year) %>%
  tally %>%
  ungroup %>%
  filter(year >= 2015) %>%
  group_by(year) %>%
  summarise(`Total Observations` = sum(n))
tablify(x)
```

# Combined

## Number of students that we have both info on absenteeism and performance

```{r}
x <- 
  ab %>%
  group_by(name, year) %>%
  tally %>%
  ungroup %>%
  filter(year >= 2015) %>% 
  mutate(src = 'ab') %>%
  bind_rows(
    performance %>%
  group_by(name, year) %>%
  tally %>%
  ungroup %>%
  filter(year >= 2015) %>%
    mutate(src = 'perf')
  ) %>%
  group_by(year, name, src) %>%
  summarise(n = sum(n)) %>%
  ungroup %>%
  arrange(name, year) %>%
  group_by(year, name) %>%
  summarise(have_ab = sum(n[src == 'ab'] > 0),
            have_perf = sum(n[src == 'perf'] > 0)) %>%
  mutate(have_both = have_ab == 1 & have_perf == 1) %>%
  group_by(year) %>%
  summarise(`Students in absenteeism` = sum(have_ab),
            `Students in performance` = sum(have_perf),
            `Students in both` = sum(have_both)) 
tablify(x)
```

## Total magnitude of database

- `r nrow(ab)` student-days observed (absenteeism)  

- `r nrow(performance)` student-class-trimesters observed (performance)

## Average absenteeism rate, per term, per school

```{r}
x <- 
  ab %>%
  group_by(year_term, school, district) %>%
  summarise(absences = length(which(absent)),
            presences = length(which(!absent)),
            eligible_days = n(),
            Students = length(unique(name))) %>%
  mutate(absenteeism_rate = absences / eligible_days * 100) %>%
  ungroup %>%
  filter(!grepl('NA', year_term))
y <- x
names(y) <- gsub('_', ' ', names(y))
names(y) <- Hmisc::capitalize(names(y))
# tablify(y)
```

```{r}
library(RColorBrewer)
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Set1'))(length(unique(x$school)))
ggplot(data = x,
       aes(x = year_term,
           y = absenteeism_rate,
           group = school,
           color = school)) +
  geom_line(alpha = 0.7,
            size = 2) +
  theme_cism() +
  scale_color_manual(name = 'School',
                     values = cols) +
  ylim(0, 20) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = 'Year-term',
       y = 'Absenteeism rate')

```


## Average absenteeism rate, per term, per district

```{r}
x <- 
  ab %>%
  group_by(year_term, district) %>%
  summarise(absences = length(which(absent)),
            presences = length(which(!absent)),
            eligible_days = n(),
            Students = length(unique(name))) %>%
  mutate(absenteeism_rate = absences / eligible_days * 100) %>%
  ungroup %>%
  filter(!grepl('NA', year_term))
y <- x
names(y) <- gsub('_', ' ', names(y))
names(y) <- Hmisc::capitalize(names(y))
# tablify(y)
```



```{r}
library(RColorBrewer)
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Set1'))(length(unique(x$district)))
ggplot(data = x,
       aes(x = year_term,
           y = absenteeism_rate,
           group = district,
           color = district)) +
  geom_line(alpha = 0.7,
            size = 2) +
  theme_cism() +
  scale_color_manual(name = 'District',
                     values = cols) +
  ylim(0, 10) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = 'Year-term',
       y = 'Absenteeism rate')

```

# Additional charts

## Sex

```{r}
# Sex and absenteeism
x <- ab %>%
  left_join(census %>%
              dplyr::select(census_name, sex)) %>%
  group_by(sex) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  filter(!is.na(sex))

ggplot(data = x,
       aes(x = sex,
           y = absenteeism_rate)) +
  geom_bar(stat = 'identity',
           fill = 'darkgreen',
           alpha = 0.6) +
  theme_cism() +
  labs(x = 'Sex',
       y = 'Absenteeism rate',
       title = 'Absenteeism rate by sex')
```


## Sex and absetneeism over time

```{r}
x <- ab %>%
  left_join(census %>%
              dplyr::select(census_name, sex)) %>%
  mutate(year_month = date_truncate(date, 'month')) %>%
  group_by(sex, year_month, district) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  filter(!is.na(sex))

ggplot(data = x,
       aes(x = year_month,
           y = absenteeism_rate,
           color = district)) +
  geom_line() + 
  facet_wrap(~sex, ncol = 1) +
  scale_color_manual(name = 'District',
                     values = cols) +
  theme_cism() +
  labs(x = 'Month',
       y = 'Absenteeism rate',
       title = 'Absenteeism rate by sex')
```

## Distance to school

```{r}

# DISTANCE TO SCHOOL AND ABSENTEEISM
x <- 
  ab %>%
  group_by(census_name) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  left_join(students %>%
              dplyr::select(census_name,
                            km_to_health_facility,
                            km_to_school,
                            km_to_other_district_student)) %>%
  filter(!is.na(km_to_school)) %>%
  mutate(distance_to_school = 
           ifelse(km_to_school < 5, '0-5', '5+'))

ggplot(data = x,
       aes(x = km_to_school,
           y = absenteeism_rate)) +
  geom_jitter(alpha = 0.2, color = 'darkorange') +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = 'Kilometers to school',
       y = 'Absenteeism rate',
       title = 'Absenteeism and distance to school',
       subtitle = 'Closer = better') +
  theme_cism() +
  geom_smooth()
```

## Distance to school

```{r}
ggplot(data = x,
       aes(x = km_to_school,
           y = absenteeism_rate)) +
  geom_jitter(alpha = 0.2, color = 'darkorange') +
  # scale_x_log10() +
  # scale_y_log10() +
  labs(x = 'Kilometers to school',
       y = 'Absenteeism rate',
       title = 'Absenteeism and distance to school',
       subtitle = 'Closer = better') +
  theme_cism() +
  geom_smooth()
```

## Distance to school

```{r}

ggplot(data = x,
       aes(x = km_to_school,
           y = absenteeism_rate)) +
  geom_jitter(alpha = 0.2, color = 'darkorange') +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = 'Kilometers to school',
       y = 'Absenteeism rate',
       title = 'Absenteeism and distance to school',
       subtitle = 'Closer = better') +
  theme_cism() +
  geom_smooth() +
  facet_wrap(~distance_to_school,
             ncol = 1,
             scales = 'free_x')

```

## Number of siblings

```{r}

# Number of siblings and absetneeism rate

x <- ab %>%
  group_by(census_name) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  left_join(census %>%
              dplyr::select(census_name, n_children_agregado)) %>%
  mutate(siblings = ifelse(n_children_agregado == 1, '0',
                           ifelse(n_children_agregado <=3, '1-2',
                                  ifelse(n_children_agregado <=6, '3-5',
                                         ifelse(n_children_agregado >=5, '6+', NA))))) %>%
  filter(!is.na(siblings)) %>%
  mutate(n_children = ifelse(siblings <= 2, '1-3',
                                    '4+'))

avg <- x %>%
  group_by(siblings) %>%
  summarise(absenteeism_rate = mean(absenteeism_rate))

ggplot(data = x,
       aes(x = siblings,
           y = absenteeism_rate)) +
  geom_jitter(alpha = 0.6,
              color = 'darkgreen') +
  geom_violin(alpha = 0.3) +
  geom_point(data = avg,
           aes(x = siblings, 
               y = absenteeism_rate),
           color = 'red',
           size = 40,
           pch = '-') +
  scale_y_log10() +
  theme_cism() +
  labs(x = 'Siblings',
       y = 'Absenteeism rate',
       title = 'Absenteeism rate by siblings',
       subtitle = 'No clear relationship')
```

## Number of siblings

```{r}


x <- ab %>%
  group_by(census_name, year_term, district) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  left_join(census %>%
              dplyr::select(census_name, n_children_agregado)) %>%
  mutate(siblings = ifelse(n_children_agregado == 1, '0',
                           ifelse(n_children_agregado <=3, '1-2',
                                  ifelse(n_children_agregado <=6, '3-5',
                                         ifelse(n_children_agregado >=5, '6+', NA))))) %>%
  filter(!is.na(siblings)) %>%
  mutate(n_children = ifelse(siblings <= 2, '1-3',
                                    '4+')) %>%
  group_by(district, year_term, n_children) %>%
  summarise(absenteeism_rate = mean(absenteeism_rate)) %>%
  filter(!grepl('NA', year_term))

ggplot(data = x,
       aes(x = year_term,
           y = absenteeism_rate,
           color = district,
           group = district)) +
  geom_line() + 
  facet_wrap(~n_children, ncol = 1) +
  labs(x = 'Year/term',
       y = 'Absenteeism rate',
       title = 'Effect of family size',
       subtitle = 'Absenteeism rate by number of children in agregado') +
  theme_cism() +
  scale_color_manual(name = 'District',
                     values = cols) +
  theme(axis.text.x = element_text(angle = 90))
```

## Education

```{r}
x <- ab %>%
  left_join(census %>%
              dplyr::select(census_name, ses_education_household))  %>%
  group_by(year_term, census_name, district) %>%
  summarise(absences = length(which(absent)),
            eligibles = n(),
            ses_education_household = first(ses_education_household)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  filter(!is.na(ses_education_household)) %>%
  filter(!grepl('NA', year_term)) %>%
  mutate(education = ifelse(ses_education_household <2, 
                            'Low',
                            ifelse(ses_education_household <4,
                                   'Medium', 
                                   ifelse(ses_education_household >= 4, 'High', NA)))) %>%
  mutate(education = factor(education,
                            levels = c('Low', 'Medium', 'High')))
         

ggplot(data = x,
       aes(x = ses_education_household,
           y = absenteeism_rate)) +
  geom_jitter(width = 1,
              alpha = 0.3,
              color = 'darkgreen') +
  geom_smooth() +
  theme_cism() +
  labs(x = 'Household education level',
       y = 'Absenteeism rate',
       title = 'Education and absenteeism',
       subtitle = 'No clear relationship')

```

## Education and absenteeism

```{r}
ggplot(data = x,
       aes(x = education,
           y = absenteeism_rate)) +
  geom_jitter(#width = 1,
              alpha = 0.3,
              color = 'darkgreen') +
  geom_violin() +
  theme_cism() +
  labs(x = 'Household education level',
       y = 'Absenteeism rate',
       title = 'Education and absenteeism')
```

## Education and absenteeism

```{r}


ggplot(data = x,
       aes(x = absenteeism_rate,
           group = education,
           fill = education)) +
  geom_density(alpha = 0.8, n = 15) +
  scale_x_log10() +
  labs(x = 'Absenteeism rate',
       y = 'Density',
       title = 'Family education and absenteeism') +
  scale_fill_manual(name = 'Education',
                    values = brewer.pal(3, 'Spectral')) +
  theme_cism()
```

## Education and absenteeism over time

```{r}
# Over time
x <- x %>%
  group_by(year_term,
           education, district) %>%
  summarise(absenteeism_rate = sum(absences) / sum(eligibles) * 100) %>%
  ungroup
ggplot(data = x,
       aes(x = year_term,
           y = absenteeism_rate,
           group = district,
           color = district)) +
  geom_line() +
  geom_point() +
  facet_wrap(~education, ncol = 3) +
  theme_cism() +
  labs(x = 'Term',
       y = 'Absenteeism rate',
       title = 'Effect of Education level') +
  scale_color_manual(name = 'District',
                     values = cols) +
  theme(axis.text.x = element_text(angle = 90))
```


## SES Assets

```{r}
x <- ab %>%
  left_join(census %>%
              dplyr::select(census_name, ses_asset_score,
                            ses_conditions_score))  %>%
  group_by(year_term, census_name, district) %>%
  summarise(absences = length(which(absent)),
            eligibles = n(),
            ses_asset_score = first(ses_asset_score),
            ses_conditions_score = first(ses_conditions_score)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  # filter(!is.na(ses_education_household)) %>%
  filter(!grepl('NA', year_term)) %>%
  mutate(conditions = ifelse(ses_conditions_score <= 10000, 
                             'Low',
                             ifelse(ses_conditions_score <= 2000,
                                    'Medium',
                                    ifelse(ses_conditions_score > 20000,
                                           'High', NA)))) %>%
  mutate(assets = ifelse(ses_asset_score <= 6780, 
                         'Low',
                         ifelse(ses_asset_score > 6780,
                                'High', 
                                NA))) %>%
  filter(!is.na(assets),
         !is.na(conditions))


ggplot(data = x,
       aes(x = assets,
           y = absenteeism_rate)) +
  geom_jitter(color = 'darkgreen', alpha = 0.6) +
  geom_violin(alpha = 0.3) +
  scale_y_log10() +
  labs(x = 'Assets',
       y = 'Absenteeism rate',
       title = 'Assets and absenteeism',
       subtitle = 'No finding') +
  theme_cism()
```

## Assets

```{r}

y <- x %>%
  group_by(year_term,
           district,
           assets) %>%
  summarise(absenteeism_rate = sum(absences) / sum(eligibles) * 100)

ggplot(data = y,
       aes(x = year_term,
           y = absenteeism_rate,
           color = district,
           group = district)) +
  geom_line() +
  facet_wrap(~assets, ncol = 1) +
  # geom_smooth() +
  theme_cism() +
  labs(x = 'Term',
       y = 'Absenteeism rate',
       title = 'Assets score and absenteeism') +
  scale_color_manual(name = 'District',
                     values = cols) +
  theme(axis.text.x = element_text(angle = 90))
```

## Assets

```{r}

ggplot(data = x,
       aes(x = conditions,
           y = absenteeism_rate)) +
  geom_jitter(color = 'darkgreen', alpha = 0.6) +
  geom_violin(alpha = 0.3) +
  scale_y_log10() +
  labs(x = 'Conditions score',
       y = 'Absenteeism rate',
       title = 'Conditions and absenteeism',
       subtitle = 'No finding') +
  theme_cism()
```

## Assets


```{r}
ggplot(data = x,
       aes(x = assets,
           y = absenteeism_rate)) +
  geom_jitter(color = 'darkgreen', alpha = 0.6) +
  geom_violin(alpha = 0.3) +
  scale_y_log10() +
  labs(x = 'Assets',
       y = 'Absenteeism rate',
       title = 'Assets and absenteeism',
       subtitle = 'No finding') +
  theme_cism()
```

## Conditions

```{r}

y <- x %>%
  group_by(year_term,
           district,
           conditions) %>%
  summarise(absenteeism_rate = sum(absences) / sum(eligibles) * 100)

ggplot(data = y,
       aes(x = year_term,
           y = absenteeism_rate,
           color = district,
           group = district)) +
  geom_line() +
  facet_wrap(~conditions, ncol = 1) +
  # geom_smooth() +
  theme_cism() +
  labs(x = 'Term',
       y = 'Absenteeism rate',
       title = 'Conditions score and absenteeism') +
  scale_color_manual(name = 'District',
                     values = cols) +
  theme(axis.text.x = element_text(angle = 90))
```

## Performance over time

```{r}

# See how performance changed over time
x <- performance %>% 
  group_by(district,year, trimester) %>% 
  summarise(average=mean(value, na.rm=TRUE),
            students = length(unique(`Study Subject ID`))) %>%
  mutate(year_trimester = paste0(year, '-', trimester)) %>%
  filter(!is.na(year))
#
ggplot(data = x,
       aes(x = year_trimester,
           y = average,
           group = district,
           color = district)) +
  geom_point(aes(size = students)) +
  geom_line()
```

## Performance over time

```{r}

# See how performance changed over year
x <- performance %>% 
  group_by(district,year) %>% 
  summarise(average=mean(value, na.rm=TRUE),
            students = length(unique(`Study Subject ID`))) %>%
  filter(!is.na(year))
#
ggplot(data = x,
       aes(x = year,
           y = average,
           group = district,
           color = district)) +
  geom_point(aes(size = students)) +
  geom_line()
```


## Performance over time

```{r}

# See how performance changed over time by school
x <- performance %>% 
  group_by(district, school,year, trimester) %>% 
  summarise(average=mean(value, na.rm=TRUE),
            students = length(unique(`Study Subject ID`))) %>%
  mutate(year_trimester = paste0(year, '-', trimester)) %>%
  filter(!is.na(year))
#
library(RColorBrewer)
cols <- colorRampPalette(brewer.pal(9, 'Spectral'))(length(unique(x$school)))
ggplot(data = x,
       aes(x = year_trimester,
           y = average,
           group = school,
           color = school)) +
  geom_point(aes(size = students,
                 pch = district),
             alpha = 0.8) +
  geom_line() +
  scale_color_manual(name = 'School',
                     values = cols) +
  theme_bw()
```


## Performance over time

```{r}



# Distribution of performance by year and district
ggplot(data = performance,
       aes(x = value)) +
  geom_histogram(alpha = 0.5) +
  facet_wrap(~year, ncol = 1)

```


## Performance over time

```{r}

ggplot(data = performance,
       aes(x = value)) +
  geom_histogram(alpha = 0.5) +
  facet_grid(year~district)

```


## Performance over time

```{r}

ggplot(data = performance %>%
         mutate(year = factor(year)),
       aes(x = value,
           group = year,
           fill = year)) +
  geom_density(alpha = 0.5,
               color = NA) +
  facet_wrap(~school)

```


## Performance over time

```{r}

ggplot(data = performance %>%
         mutate(year = factor(year)),
       aes(x = year,
           y = value)) +
  # geom_point(alpha = 0.6) +
  geom_violin(alpha = 0.3, 
              fill = 'darkorange') +
  facet_wrap(~school)

```

# Other indictors

## Monthly absenteeism by district

```{r}
monthly_by_district <-
  ab %>%
  group_by(district, month) %>%
  summarise(absences = length(which(absent)),
            presences = length(which(!absent)),
            eligible_days = n()) %>%
  mutate(absenteeism_rate = absences / eligible_days * 100) %>%
  ungroup

ggplot(data = monthly_by_district,
       aes(x = month,
           y = absenteeism_rate,
           group = district,
           color = district)) +
  geom_line(alpha = 0.8,
            size = 2) +
  theme_cism() +
  labs(x = 'Date',
       y = 'Absenteeism rate',
       title = 'Montly absenteeism rate',
       subtitle = 'One district = one line') +
  scale_color_manual(name = 'District',
                     values = cols)
```

## Absenteeism by term and district

```{r}
# year term by district
year_term_by_district <-
  ab %>%
  group_by(year_term, district) %>%
  summarise(absences = length(which(absent)),
            presences = length(which(!absent)),
            eligible_days = n()) %>%
  mutate(absenteeism_rate = absences / eligible_days * 100) %>%
  ungroup %>%
  filter(!is.na(year_term)) %>%
  filter(!grepl('NA', year_term))

ggplot(data = year_term_by_district,
       aes(x = year_term,
           y = absenteeism_rate,
           group = district,
           color = district)) +
  geom_line(alpha = 0.8,
            size = 2) +
  theme_cism() +
  labs(x = 'Year - term',
       y = 'Absenteeism rate',
       title = 'Montly absenteeism rate',
       subtitle = 'One district = one line') +
  scale_color_manual(name = 'District',
                     values = cols)

```

## Absenteeism by year and district

```{r}
year_by_district <-
  ab %>%
  group_by(year, district) %>%
  summarise(absences = length(which(absent)),
            presences = length(which(!absent)),
            eligible_days = n()) %>%
  mutate(absenteeism_rate = absences / eligible_days * 100) %>%
  ungroup

library(ggrepel)
ggplot(data = year_by_district,
       aes(x = factor(year),
           y = absenteeism_rate,
           group = district,
           color = district)) +
  geom_line(alpha = 0.8,
            size = 2) +
  geom_point(size = 5) +
  theme_cism() +
  labs(x = 'Year - term',
       y = 'Absenteeism rate',
       title = 'Montly absenteeism rate',
       subtitle = 'One district = one line') +
  scale_color_manual(name = 'District',
                     values = cols) +
  geom_hline(yintercept = year_by_district %>%
               filter(year == min(year)) %>%
               .$absenteeism_rate,
             lty = 2,
             alpha = 0.6) +
  geom_text_repel(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')),
                  point.padding = unit(1, "lines")) +
  ylim(0, 8)
```

## Standardized absenteeism


```{r}

# Standardized
year_by_district <-
  year_by_district %>%
  arrange(year) %>%
  group_by(district) %>%
  mutate(p = absenteeism_rate / first(absenteeism_rate) * 100)

ggplot(data = year_by_district,
       aes(x = factor(year),
           y = p,
           group = district,
           color = district)) +
  geom_line(alpha = 0.8,
            size = 2) +
  geom_point(size = 5) +
  theme_cism() +
  labs(x = 'Year - term',
       y = 'Absenteeism rate',
       title = 'Montly absenteeism rate',
       subtitle = 'One district = one line') +
  scale_color_manual(name = 'District',
                     values = cols) +
  geom_hline(yintercept = 100,
             lty = 2,
             alpha = 0.6) +
  geom_text_repel(aes(label = paste0(round(p, digits = 2), '%')),
                  point.padding = unit(1, "lines"))

```
